The problem and promise of scale in animal social multilayer networks

### 2020-02-07 paper/manuscript.Rmd
space units

## Landscape data and habitat classification
Landcover data were provided by the Newfoundland and Labrador Wildlife Division.
Available landcover classification included nine habitat types at <!-- TODO:
double check lc res--> 10m resolution. We reclassified the landcover types into
three categories: foraging habitat, open habitat, and forest habitat using the R
package `raster` [@Hijmans_2019]. Foraging habitat <!-- TODO: "foraging" should
be "lichen" ?--> consisted of lichen habitat, open habitat consisted of wetland,
rocky barrens, and anthropogenic habitat types, while forest habitat consisted
of coniferous forest, conifer scrub, broadleaf forest, and mixed-wood forest
habitat types. Water habitat was excluded from all subsequent analyses.

@@

Within each layer, we assumed individuals were associating if simultaneous GPS
fixes (i.e., recorded within 5 minutes of each other) were within a given
spatial distance of one another. Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is applied for group assignment. For GPS data, the
chain rule is applied if discrete spatiotemporal GPS fixes buffered by a given
distance formed a contiguous buffer for two or more individuals, even if some
individuals within the buffer were not within the given distance of one another
[@Robitaille_2019]. Group assignment based on the chain rule has previously been
applied to caribou at a distance of 50m [@Peignier_2019; @Lesmerises_2018]. In
all networks, nodes represented individuals, intralayer edges represented
associations between individuals in a given habitat type and season, and
interlayer edges represented connections between the same individuals across
contexts.

@@

We altered the social scale based on visual (short distance) and auditory (long
distance) sensory modalities of caribou<!-- TODO: REF?-->. Specifically, we
modified the spatial distance threshold required for group assignment using
`spatsoc` [@Robitaille_2019] (see above) and re-assigned groups at incremental
distances between 5-500m. For example, at the finest social scale, only
individuals within 5m of one another were considered in the same group, whereas
at the coarsest social scale, individuals within 500m of one another were
considered in the same group.

@@

To assess the influence of spatial scale on multilayer networks, we aggregated
our landcover raster at varying scales relevant to caribou ecology.
We aggregated the 10m landcover raster using the R package `grainchanger`
[@Graham_2019] at five scales: 100m, 250m, 500m, and 1000m. Aggregation occurred
using a modal moving window method using a circular window corresponding to the
above scales [@Graham_2019a]. 10m and 100m represents fine-scale decision making
for caribou during foraging (Webber et al. unpublished data), while re-sampling at 500m and 1000m
represents the scale at which caribou tend to select and avoid habitat
(Bastille-Rousseau et al. 2017).

@@

<!-- TODO: discuss layer similarity for social threshold -->
In general, individuals in multilayer networks became more connected as social scale 
increased; however, this relationship varied by habitat type and
season. As social scale increased, connective redundancy increased rapidly when
groups were formed up to approximately 100m threshold, followed by a plateau and slight
decrease at higher thresholds, presumably due to an increase in summer forest
degree at ~500m (Figure \@ref(fig:socres)d). Regardless of social scale, layer
relevance did not change for any habitat types or seasons, which suggests
habitat-by-season layers captured most, if not all, of the potential connections
an individual would have across all networks (Figure \@ref(fig:socres)b).
Finally, somewhat predictably, graph strength for winter networks increased up
to ~100-200m thresholds, followed by a plateau, while graph strength for summer
networks did not change as the social threshold increased (Figure
\@ref(fig:socres)a).  Taken together, these results suggest the optimal social
scale at which groups should be assigned is likely somewhere between ~20-100m
based on the rate at which network metrics stabilized. For ungulates, groups are
often assigned based on 50m threshold (Lingle 2003; Lesmeries et al. 2018;
Peignier et al. 2019), suggesting that given similar quantities of data, ~50m is
likely an ideal social scale for generating social networks in caribou and
similar species.

<!-- landcover resolution
* decreasing availability of patchy lichen = no edges in these layers, though some small proportion of relocations still exist in those layers
* decreasing connective redundancy because through resolution because lichen decreased in connectivity and landcovers became more homogeneous (?)
* decreasing winter lichen relevance because the degree + number of edges decreased
--> 
<!-- TODO: discuss things in a,b,c,d order consistently. They shouldn't be presented out of order -->
Increasing the scale of landcover resolution resulted in decreased proportion of rare and
patchy habitats and low connectivity of individuals in these habitats. In
winter, layer relevance was stable for open and forest habitats but decreased
for forage habitats (Figure \@ref(fig:lcres)b). Connective redundancy decreased
with increasing landcover resolution due the decrease in connectivity of
individuals in winter forage habitats, resulting in a multilayer network where
neighbours were less frequently repeated in multiple layers <!-- TODO: I'm having a hard 
time understanding this previous sentence --> (Figure\@ref(fig:lcres)d). Layer similarity 
was generally stable for forest and open
habitats except for a contrasting increase in similarity for forest and decrease
in similarity for open habitats at 250m landcover resolution (Figure
\@ref(fig:lcres)c). Layer similarity was not possible to calculate between
winter and summer forage because, at all landcover resolutions greater than 10m
there were no individuals connected in the network layers. <!-- individuals
still used forage but in much lower proportion + lowered opportunity to
associate --> These results indicate the importance of matching landcover
resolution to scale of selection as well as ensuring the landcover product used
has a resolution fine enough to detect relevant landscape features and habitats. <!-- TODO: how? This isn't clear enough for me to follow yet -->

@@

(ref:socres) Varying scale in social threshold. For each social threshold (5m, 50m, 100m, 150m, 200m, 250m, 300m, 350m, 400m, 450m, 500m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter) and four network metrics calculated.  A) Mean graph strength across individuals showed an increase for winter network layers at 50m through 500m while summer network layers had consistently weak connections. B) Mean layer relevance was relatively stable across social thresholds for each season and landcover layer. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an overall increase in similarity for open layers and decrease in similarity for lichen layers. <!--TODO: what about forest.--> D) Mean connective redundancy across individuals increased rapidly between 5m and 50m reaching a plateau  of about 0.65 at approximately a 50m <!--TODO:**??**--> social threshold before decreasing between 400m and 500m.

@@

(ref:lcres) Variation in multilayer network metrics as a function of changes in landcover resolution, a metric which we interpret as spatial scale. For each landcover resolution (10m, 100m, 250m, 500m and 1000m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter), and four network metrics were calculated and we examined how network metrics changed as a function of changes in spatial scale. A) Mean graph strength of individuals was weaker during the summer compared to the winter months. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter lichen layer relevance at higher landcover resolutions. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an increase in similarity for forest layers at 250m and decrease in similarity for open layers at 250m. Note that lichen layer similarity is not shown because individuals did not interact in lichen habitats greater greater than 10m landcover resolution. D) Mean connective redundancy across individuals showed a decreasing trend with increasing landcover resolution.

>>>>

## Landscape data and habitat classification
Landcover data were provided by the Newfoundland and Labrador Wildlife Division.
Available landcover classification included nine habitat types at <!-- TODO:
double check lc res--> 10 m resolution. We reclassified the landcover types into
three categories: foraging habitat, open habitat, and forest habitat using the R
package `raster` [@Hijmans_2019]. Foraging habitat <!-- TODO: "foraging" should
be "lichen" ?--> consisted of lichen habitat, open habitat consisted of wetland,
rocky barrens, and anthropogenic habitat types, while forest habitat consisted
of coniferous forest, conifer scrub, broadleaf forest, and mixed-wood forest
habitat types. Water habitat was excluded from all subsequent analyses.

@@

Within each layer, we assumed individuals were associating if simultaneous GPS
fixes (i.e., recorded within 5 minutes of each other) were within a given
spatial distance of one another. Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is applied for group assignment. For GPS data, the
chain rule is applied if discrete spatiotemporal GPS fixes buffered by a given
distance formed a contiguous buffer for two or more individuals, even if some
individuals within the buffer were not within the given distance of one another
[@Robitaille_2019]. Group assignment based on the chain rule has previously been
applied to caribou at a distance of 50 m [@Peignier_2019; @Lesmerises_2018]. In
all networks, nodes represented individuals, intralayer edges represented
associations between individuals in a given habitat type and season, and
interlayer edges represented connections between the same individuals across
contexts.

@@

We altered the social scale based on visual (short distance) and auditory (long
distance) sensory modalities of caribou<!-- TODO: REF?-->. Specifically, we
modified the spatial distance threshold required for group assignment using
`spatsoc` [@Robitaille_2019] (see above) and re-assigned groups at incremental
distances between 5-500 m. For example, at the finest social scale, only
individuals within 5 m of one another were considered in the same group, whereas
at the coarsest social scale, individuals within 500 m of one another were
considered in the same group.

@@

To assess the influence of spatial scale on multilayer networks, we aggregated
our landcover raster at varying scales relevant to caribou ecology.
We aggregated the 10 m landcover raster using the R package `grainchanger`
[@Graham_2019] at five scales: 100 m, 250 m, 500 m, and 1000 m. Aggregation occurred
using a modal moving window method using a circular window corresponding to the
above scales [@Graham_2019a]. 10 m and 100 m represents fine-scale decision making
for caribou during foraging (Webber et al. unpublished data), while re-sampling at 500 m and 1000 m
represents the scale at which caribou tend to select and avoid habitat
(Bastille-Rousseau et al. 2017).

@@

<!-- TODO: discuss layer similarity for social threshold -->
In general, individuals in multilayer networks became more connected as social scale 
increased; however, this relationship varied by habitat type and
season. As social scale increased, connective redundancy increased rapidly when
groups were formed up to approximately 100 m threshold, followed by a plateau and slight
decrease at higher thresholds, presumably due to an increase in summer forest
degree at ~500 m (Figure \@ref(fig:socres)d). Regardless of social scale, layer
relevance did not change for any habitat types or seasons, which suggests
habitat-by-season layers captured most, if not all, of the potential connections
an individual would have across all networks (Figure \@ref(fig:socres)b).
Finally, somewhat predictably, graph strength for winter networks increased up
to ~100-200 m thresholds, followed by a plateau, while graph strength for summer
networks did not change as the social threshold increased (Figure
\@ref(fig:socres)a).  Taken together, these results suggest the optimal social
scale at which groups should be assigned is likely somewhere between ~20-100 m
based on the rate at which network metrics stabilized. For ungulates, groups are
often assigned based on 50 m threshold (Lingle 2003; Lesmeries et al. 2018;
Peignier et al. 2019), suggesting that given similar quantities of data, ~50 m is
likely an ideal social scale for generating social networks in caribou and
similar species.

<!-- landcover resolution
* decreasing availability of patchy lichen = no edges in these layers, though some small proportion of relocations still exist in those layers
* decreasing connective redundancy because through resolution because lichen decreased in connectivity and landcovers became more homogeneous (?)
* decreasing winter lichen relevance because the degree + number of edges decreased
--> 
<!-- TODO: discuss things in a,b,c,d order consistently. They shouldn't be presented out of order -->
Increasing the scale of landcover resolution resulted in decreased proportion of rare and
patchy habitats and low connectivity of individuals in these habitats. In
winter, layer relevance was stable for open and forest habitats but decreased
for forage habitats (Figure \@ref(fig:lcres)b). Connective redundancy decreased
with increasing landcover resolution due the decrease in connectivity of
individuals in winter forage habitats, resulting in a multilayer network where
neighbours were less frequently repeated in multiple layers <!-- TODO: I'm having a hard 
time understanding this previous sentence --> (Figure\@ref(fig:lcres)d). Layer similarity 
was generally stable for forest and open
habitats except for a contrasting increase in similarity for forest and decrease
in similarity for open habitats at 250 m landcover resolution (Figure
\@ref(fig:lcres)c). Layer similarity was not possible to calculate between
winter and summer forage because, at all landcover resolutions greater than 10 m
there were no individuals connected in the network layers. <!-- individuals
still used forage but in much lower proportion + lowered opportunity to
associate --> These results indicate the importance of matching landcover
resolution to scale of selection as well as ensuring the landcover product used
has a resolution fine enough to detect relevant landscape features and habitats. <!-- TODO: how? This isn't clear enough for me to follow yet -->

@@

(ref:socres) Varying scale in social threshold. For each social threshold (5 m, 50 m, 100 m, 150 m, 200 m, 250 m, 300 m, 350 m, 400 m, 450 m, 500 m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter) and four network metrics calculated.  A) Mean graph strength across individuals showed an increase for winter network layers at 50 m through 500 m while summer network layers had consistently weak connections. B) Mean layer relevance was relatively stable across social thresholds for each season and landcover layer. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an overall increase in similarity for open layers and decrease in similarity for lichen layers. <!--TODO: what about forest.--> D) Mean connective redundancy across individuals increased rapidly between 5 m and 50 m reaching a plateau  of about 0.65 at approximately a 50 m <!--TODO:**??**--> social threshold before decreasing between 400 m and 500 m.

@@

(ref:lcres) Variation in multilayer network metrics as a function of changes in landcover resolution, a metric which we interpret as spatial scale. For each landcover resolution (10 m, 100 m, 250 m, 500 m and 1000 m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter), and four network metrics were calculated and we examined how network metrics changed as a function of changes in spatial scale. A) Mean graph strength of individuals was weaker during the summer compared to the winter months. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter lichen layer relevance at higher landcover resolutions. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an increase in similarity for forest layers at 250 m and decrease in similarity for open layers at 250 m. Note that lichen layer similarity is not shown because individuals did not interact in lichen habitats greater greater than 10 m landcover resolution. D) Mean connective redundancy across individuals showed a decreasing trend with increasing landcover resolution.
